---
title: "Installing ITKR and ANTsR"
author: "John Muschelli"
date: "August 29, 2016"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
---

# Preface

The course is designed for minimal system requirements or installation knowledge.  We highly encourage users to install the [VirtualBox](https://www.smart-stats-tools.org/mooc-2015)

# Windows Setup

FSL will not install on a Windows machine, and the developers of FSL recommend a virtual machine to their users: [http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FslInstallation/Windows](http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FslInstallation/Windows).

# Mac OSX Setup

## Installing Xcode
In order to install packages with compilation in `R`, you must install `Xcode`.  For example, if you get the error: "Error: Could not find build tools necessary to build" for a package, please see [this StackOverflow Post](http://stackoverflow.com/questions/35096233/error-could-not-find-build-tools-necessary-to-build-dplyr) for a discussion and [Xcode from Apple](https://developer.apple.com/downloads/).  

## Installing Git and CMake

You need `cmake` and `git` to install `ITKR`.  Go to a `Terminal` (using spotlight or from `/Applications/Utilities/Terminal`) and type

```
which cmake
which git
```

### Installing Git

If nothing comes up for `git`, please see [https://www.atlassian.com/git/tutorials/install-git/mac-os-x](https://www.atlassian.com/git/tutorials/install-git/mac-os-x) to install `git`.

### Installing CMake

CMake can be installed from [https://cmake.org/download/](https://cmake.org/download/) . 

#### Installing CMake from Mac DMG (recommended)

If you download the CMake dmg, you must add the cmake to your PATH, by adding:

```
PATH=$PATH:/Applications/CMake.app/Contents/bin/
export PATH
```

the the end of your `~/.bash_profile` file on your machine.

You can edit `~/.bash_profile` by opening a `Terminal` (again `/Applications/Utilities/Terminal`) and typing:
```
vi ~/.bash_profile
```

Type the letter `i` so that the bottom says `INSERT`, go to the bottom, and paste the command from above with `PATH`.  When this is done, type `Esc + : + wq`. So that it says `:wq` at the bottom (stands for `w`rite out the file again, and then `q`uit).

#### Installing CMake from Source 

You can also build Cmake from source: 

Download using: [https://cmake.org/files/v3.6/cmake-3.6.1.tar.gz](https://cmake.org/files/v3.6/cmake-3.6.1.tar.gz) (or the latest on their website).  Double click the `.tar.gz` file to unzip it.

The README.rst file are instructions on how to build cmake.  If you open a `Terminal` (again `/Applications/Utilities/Terminal`), change directories into the directory you unzipped the `.tar.gz`, usually by:
```
cd ~/Downloads/cmake-3.6.1
```

and then run the following command:
```
./bootstrap && make && sudo make install
```
and type your password.

# Linux Setup

## Installing Git and CMake

You need `cmake` and `git` to install `ITKR`.  Go to a `Terminal` and type

```
which cmake
which git
```
to see if `git` and `cmake` are installed.

### Apt-get/Ubuntu/Debian

```
sudo apt-get install cmake
sudo apt-get install git-all
```

### Yum/Centos/RedHat

```
sudo yum install cmake
sudo yum install git-all
```

## `fslr` setup 

### RStudio vs. Terminal

If you run `R` from the command line/Terminal then `FSLDIR` should be set when you installed `FSL` and you should be able to run `fslr` out of the box.

If you are using `RStudio` or an `R` GUI, then you may have to make some changes to your directories for `fslr` to work.

### Neurodebian `fslr` Setup

If you install `FSL` (version 5.0) using Neurodebian, the following script should copy the necessary data/binaries to a path where `fslr` can find them.  

```{bash, eval = FALSE}
FSLDIR=/usr/local/fsl
FSLSHARE=/usr/share/data
PATH=$PATH:${FSLDIR}/bin

sudo mkdir -p ${FSLDIR}/bin
sudo cp /usr/lib/fsl/5.0/* ${FSLDIR}/bin/
sudo mkdir -p ${FSLDIR}/data/standard 
sudo mkdir -p ${FSLDIR}/data/atlases 

sudo cp -R ${FSLSHARE}/fsl-mni152-templates/* ${FSLDIR}/data/standard/

# setting up atlases
sudo cp -R ${FSLSHARE}/harvard-oxford-atlases/* ${FSLDIR}/data/atlases/ 
sudo cp -R ${FSLSHARE}/juelich-histological-atlas/* ${FSLDIR}/data/atlases/ 
sudo cp -R ${FSLSHARE}/bangor-cerebellar-atlas/* ${FSLDIR}/data/atlases/ 
sudo cp -R ${FSLSHARE}/jhu-dti-whitematter-atlas/* ${FSLDIR}/data/atlases/ 
sudo cp -R ${FSLSHARE}/forstmann-subthalamic-nucleus-atlas/* ${FSLDIR}/data/atlases/ 
sudo cp -R ${FSLSHARE}/fsl-resting-connectivity-parcellation-atlases/* ${FSLDIR}/data/atlases/ 
sudo cp -R ${FSLSHARE}/mni-structural-atlas/* ${FSLDIR}/data/atlases/ 
sudo cp -R ${FSLSHARE}/oxford-thalamic-connectivity-atlas/* ${FSLDIR}/data/atlases/ 
sudo cp -R ${FSLSHARE}/talairach-daemon-atlas/* ${FSLDIR}/data/atlases/ 
```

### Standard FSL installation Setup

If `FSL` was installed using the install scripts from the FSL website, then all files necessary should be located in `/usr/local/fsl`, and `fslr` should work out of the box.

# Installing R Packages

## Installing Packages from CRAN

The `oro.nifti`, `oro.dicom`, and `fslr` packages are all on CRAN and can be installed using the standard `install.packages` way:

```{r, eval = FALSE}
if (!require(oro.nifti)) {
  install.packages("oro.nifti")
}
if (!require(oro.dicom)) {
  install.packages("oro.dicom")
}
if (!require(fslr)) {
  install.packages("fslr")
}
```

## Installing GitHub Packages


Some of the packages used are not on CRAN, and must be installed using `devtools`.  
Note below, `fslr` is included.  If you would like to use the GitHub version (in development, aka may break), then please use the code below.  Otherwise use the CRAN version.

### Installing `devtools`

```{r github, eval = FALSE}
if (!require(devtools)) {
    install.packages("devtools")
}
```

### Installing Development version of `fslr`

```{r github2, eval = FALSE}
if (!require(fslr)) {
  devtools::install_github("muschellij2/fslr")
}
```

### Installing `ITKR`

Installing `ITKR` requires the `cmaker` package (at the time of writing):

```{r ITKR, eval = FALSE}
if (!require(cmaker)) {
  devtools::install_github("stnava/cmaker")
}
if (!require(ITKR)) {
  devtools::install_github("stnava/ITKR")
}
```

If the above doesn't work, and you have **double checked** CMake and git are installed, try the following command: 

```{r ITKR_musch, eval = FALSE}
if (!require(ITKR)) {
  devtools::install_github("muschellij2/ITKR")
}
```
This is a forked version of `ITKR` and may not be the most up to date, but has some changes to the install script

### Installing `ANTsR` and `extrantsr`

Note, if `ITKR` installation failed, then `ANTsR` installation will fail as it depends on `ITKR` and the `extrantsr` installation will fail because it depends on `ANTsR`.

```{r ANTsR, eval = FALSE}
if (!require(ANTsR)) {
  devtools::install_github("stnava/ANTsR")
}
if (!require(extrantsr)) {
  devtools::install_github("muschellij2/extrantsr")
}
```

### Installing `ANTsR` from a Binary

If `ANTsR` installation does not work above, but `ITKR` installed, you can try the release version of `ANTsR` for your respective system by downloading the tarballs from:
[https://github.com/stnava/ANTsR/releases](https://github.com/stnava/ANTsR/releases), and installing from `install.packages`:

```{r, eval = FALSE}
path_to_tarball = "~/Downloads/ANTsR_OSX_0.3.tgz" # example - must change
install.packages(path_to_tarball, repos = NULL)
```


## Updating Packages

The `ITKR` and `ANTsR` packages take a long time to install.  If you would like to update `extrantsr` and use `devtools::install_github("muschellij2/extrantsr")`, it will by default check to see if the `ITKR` and `ANTsR` packages had changed since last installation and try to upgrade them.

If you want to keep the versions of `ITKR` and `ANTsR` and just update `extrantsr`,  you can use the following code:

```{r, eval = FALSE}
devtools::install_github("muschellij2/extrantsr", upgrade_dependencies = FALSE)
```

