---
title: "Basic Data Manipulation"
author: "Bui Dinh Ngoc"
date: "June 30, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

## Read an NIfTI file
This is example to read some NIfTI files and execute some basic data manipulation, by using this example code you must be install package <b>oro.nifti,fslr,AnalyzeFMRI</b> first.

#### Download a NIfTI file from Neurohacking_data repository

```{r}
library(oro.nifti)

url <- "https://raw.githubusercontent.com/muschellij2/Neurohacking/master/Basic_Data_Manipulations/Kirby21/SUBJ0001-01-MPRAGE.nii.gz"
destfile <- "SUBJ0001-01-MPRAGE.nii.gz"
fname <- file.path(getwd(), destfile)
download.file(url, destfile,mode="wb") # NIfTI is binaryfile format

maskurl <- "https://raw.githubusercontent.com/muschellij2/Neurohacking/master/Basic_Data_Manipulations/Kirby21/SUBJ0001_mask.nii.gz"
maskdestfile <- "SUBJ0001_mask.nii.gz"
maskfname <- file.path(getwd(), maskdestfile)
download.file(maskurl, maskdestfile,mode="wb") # NIfTI is binaryfile format



T1 <- readNIfTI(fname,reorient=FALSE)

mask <- readNIfTI(maskfname, reorient=FALSE) 


```

#### Show meta data from MPRAGE file

```{r}

print (T1)

```

As you see the Dimension is : 170 x 256 x 256


#### Show meta data from MPRAGE file

```{r}

print (mask)

```

As you see the of mask is same dimension (170 x 256 x 256).


#### Visualizing orthographic T1

```{r}
orthographic(T1)

```

#### Visualizing orthographic mask

```{r}
orthographic(mask)

```

#### Visualizing orthographic T1*mask

```{r}
library(fslr) # you may need install fslr

masked.T1 <- niftiarr(T1, T1*mask)

orthographic(masked.T1)

```

#### Visualizing orthographic subtract

```{r}
library(fslr) # you may need install fslr


followurl <- "https://raw.githubusercontent.com/muschellij2/Neurohacking/master/Basic_Data_Manipulations/Kirby21/SUBJ0001-02-MPRAGE.nii.gz"
followdestfile <- "SUBJ0001-02-MPRAGE.nii.gz"
followfname <- file.path(getwd(), followdestfile)
download.file(followurl, followdestfile,mode="wb")

T1.follow <- readNIfTI(followfname, reorient=FALSE)

subtract.T1 <- niftiarr(T1, T1.follow - T1)
orthographic(subtract.T1)


```


#### Visualizing orthographic subtract process

```{r}
library(fslr) # you may need install fslr


baseurl <- "https://raw.githubusercontent.com/muschellij2/Neurohacking/master/Basic_Data_Manipulations/Kirby21/SUBJ0001-01-MPRAGE_N3.nii.gz"
basefile <- "SUBJ0001-01-MPRAGE_N3.nii.gz"
basefname <- file.path(getwd(), basefile)
download.file(baseurl, basefile,mode="wb")



followurl <- "https://raw.githubusercontent.com/muschellij2/Neurohacking/master/Basic_Data_Manipulations/Kirby21/SUBJ0001-02-MPRAGE_N3_REG.nii.gz"
followfile <- "SUBJ0001-02-MPRAGE_N3_REG.nii.gz"
followfname <- file.path(getwd(), followfile)
download.file(followurl, followfile,mode="wb")


T1.base.process <- readNIfTI(basefname, reorient=FALSE)
T1.follow.process <- readNIfTI(followfname, reorient=FALSE)

subtract.T1.process <- niftiarr(T1, T1.follow.process - T1.base.process)
orthographic(subtract.T1.process)


```

Full genarated page can be found here http://rpubs.com/ngocbd/Basic-Data-Manipulation



