---
title: "Visualizing  Images from NIfTI file"
author: "Bui Dinh Ngoc"
date: "June 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

## Read an NIfTI file
This is example to read a NIfTI file and display an image , by using this example code you must be install package <b>oro.nifti)</b> first.

#### Download a NIfTI file from Neurohacking_data repository

```{r}
library(oro.nifti)

url <- "https://raw.githubusercontent.com/muschellij2/Neurohacking_data/master/BRAINIX/NIfTI/Output_3D_File.nii.gz"
destfile <- "Output_3D_File.nii.gz"
name <- file.path(getwd(), destfile)
download.file(url, destfile,mode="wb") # NIfTI is binaryfile format
nii_T1 <- readNIfTI(destfile)


```

#### Show meta data from the file

```{r}

print (nii_T1)

```

As you see this file contain 22 images  512 x 512 pixels, one pixel using 16 bits

#### Visial an Slice an NIfTI file

```{r}
image(nii_T1,z=11,plot.type="single") 

```

#### Visualizing All Slices

```{r}

image(nii_T1) 

```

####  All Planes: Coronal, Sagittal, Axial

```{r}

orthographic(nii_T1,xyz=c(200,220,11)) 

```

#### Backmapping One Slice

```{r}
is_btw_300_400<- ((nii_T1>300) &
(nii_T1<400))
nii_T1_mask<-nii_T1
nii_T1_mask[!is_btw_300_400]=NA
overlay(nii_T1,nii_T1_mask,z=11,plot.type="
single") 

```

This example not show how to change or write NIfTI file.

