---
title: "Pre-Processing Neurohacking"
author: "Bui Dinh Ngoc"
date: "7/6/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pre-Processing

This document is demo how to use R to Pre-Processing Neuro Image .We divide image pre-processing into four main conceptual steps

1. Inhomogeneity correction
2. Spatial interpolation
3. Skull stripping
4. Spatial registration 



#### Download a NIfTI file from Neurohacking_data repository

```{r}
options(fsl.path= "/usr/share/fsl/5.0")  # fsl-complete install 2 version 4.1 and 5.0 on ubuntu/debian
library(oro.nifti)
library(fslr)

destfile <- "113-01-MPRAGE.nii.gz"
if(!file.exists(destfile))
{
  url <- "https://raw.githubusercontent.com/muschellij2/Neurohacking_data/master/kirby21/visit_1/113/113-01-MPRAGE.nii.gz"
  
  name <- file.path(getwd(), destfile)
  download.file(url, destfile,mode="wb") # NIfTI is binaryfile format
}
nim=readNIfTI("113-01-MPRAGE.nii.gz", reorient=FALSE) 

```

#### Some statistics using FSL 

```{r}
Sys.setenv("LD_LIBRARY_PATH"="/usr/local/lib/") # R process maybe ignore LD_LIBRARY_PATH so i set it in code
mean(nim)
fslstats(nim, opts= "-m")
fslstats("113-01-MPRAGE.nii.gz",opts="-m")
fslstats.help() 
```

####  Bias Field Correction Using fslr

```{r}

fast_img = fsl_biascorrect(nim, retimg=TRUE) 
orthographic(nim)
orthographic(fast_img) 

```

#### Original Image Minus N4 Bias-Corrected Image

```{r}
library(scales)
sub.bias <- niftiarr(nim, nim-fast_img)
# quantile the difference image using these as breaks
q=quantile(sub.bias[sub.bias !=0],probs = seq(0,1,by=0.1))

# get a diverging gradient palette
fcol=div_gradient_pal(low="blue",mid="yellow",high ="red")
ortho2(nim,sub.bias,col.y = alpha(fcol(seq(0,1, length=10)),
0.5), ybreaks = q, ycolorbar=TRUE, text = paste0("Original
Image Minus N4", "\n Bias-Corrected Image")) 

```

#### Compare using Histogram

```{r}
library(ggplot2)
library(reshape2)

slices = c(2, 6, 10, 14, 18)
 vals = lapply(slices, function(x) {
 cbind(img = c(nim[,,x]), fast = c(fast_img[,,x]),
 slice = x)
 })
 vals = do.call("rbind", vals)
 vals = data.frame(vals)
 vals = vals[ vals$img > 0 & vals$fast > 0, ]
 colnames(vals)[1:2] = c("Original Value", "Bias-Corrected Value")
 v = melt(vals, id.vars = "slice")
 g = ggplot(aes(x = value, colour = factor(slice)), data = v) + geom_line(stat = "density") + facet_wrap(~ variable)
 g = g + scale_colour_discrete(name = "Slice #") 
 print(g)

```

#### Brain Extraction Using fslr

```{r}

bet_fast = fslbet(infile=fast_img, retimg=TRUE) 

bet_fast_mask <- niftiarr(bet_fast, 1)
is_in_mask = bet_fast>0
bet_fast_mask[!is_in_mask]<-NA
orthographic(bet_fast)
orthographic(fast_img,bet_fast_mask) 


```

FSL's Brain Extraction Tool (BET) can be used for skull stripping
BET: fast, robust, and popular 

#### Improving Brain Segmentation 

```{r}

cog = cog(bet_fast, ceil=TRUE)
cog = paste("-c", paste(cog, collapse= " "))
bet_fast2 = fslbet(infile=fast_img,retimg=TRUE,opts=cog) 
orthographic(bet_fast2)
dim(bet_fast2)

```


#### In Summary
With FSL easy to pre-processing Neuro image and perform some task as Brain Extraction .

Full genarated file can see here : http://rpubs.com/ngocbd/Pre-Processing-Neurohacking



